 Casting multiple shadows: high-dimensional interactive data visualisation with tours and embeddings

### 2020-05-05 paper.Rmd
add tour section

Like, when using DR techniques, the tour assumes we have a real data matrix
$X$ consisting of $n$ observations in $p$ dimensions. First, the tour generates 
a  sequence of $p \times d$ orthonormal projection matrices (bases) $A_{t \in \mathbb{N}}$, where $d$ is typically $1$ or $2$. For each pair of orthormal bases $A_{t}$ and $A_{t+1}$ that are generated, the geodesic
path between them is interpolated to form intermediate frames, and giving the sense of continuous movement from one basis to another. The tour is then the continous visualisation of the  projections $Y_{t} = XA_{t}$, that is the projection of $X$ onto $A_{t]$ as the tour path is interpolated between successive bases. A *grand tour* corresponds to choosing new orthoromal bases at random; allowing a user to ascertain structure via exploring the subspace of $d$-dimensional projections. Instead of picking projections at random, a *guided tour* can be used to generate a sequence 'interesting' projections as quantified by an index function. While `liminal` is able to visualise guided tours, our focus is on the grand tour.

@@

The process of generating successive bases and interpolating between them before frames, results in the tour being a dynamic visualisation technique. Generally, if $d$ is one or two dimensions, the results are visualised as
animated density or scatter plots. 
The tour is enhanced by **interactivity**:

- play/pause the current projection
- zooming to find 'interesting' points
- highlighting sets of points

- don't need to realise the entire sequence
- instead generate new bases according to a fixed frame rate
- allows user interactions to play/pause on the current view

>>>>

Like, when using DR techniques, the tour assumes we have a real data matrix
$X$ consisting of $n$ observations in $p$ dimensions. First, the tour generates 
a  sequence of $p \times d$ orthonormal projection matrices (bases) $A_{t \in \mathbb{N}}$, where $d$ is typically $1$ or $2$. For each pair of orthormal bases $A_{t}$ and $A_{t+1}$ that are generated, the geodesic
path between them is interpolated to form intermediate frames, and giving the sense of continuous movement from one basis to another. The tour is then the continous visualisation of the  projections $Y_{t} = XA_{t}$, that is the projection of $X$ onto $A_{t}$ as the tour path is interpolated between successive bases. A *grand tour* corresponds to choosing new orthoromal bases at random; allowing a user to ascertain structure via exploring the subspace of $d$-dimensional projections. Instead of picking projections at random, a *guided tour* can be used to generate a sequence 'interesting' projections as quantified by an index function. While `liminal` is able to visualise guided tours, our focus in the case studies uses the grand tour to provide overviews of global
structure in the data.

@@

The process of generating successive bases and interpolating between them to  construct intermediary frames, means the tour is a dynamic visualisation technique. Generally, the user would set $d=2$ the tour is visualised as
a animated scatterplot. This process of constructing bases and intermediate frames and visualising the resulting projection is akin to making a "flipbook" animation. Like with a flipbook, an interface to the tour requires the ability
to interact and modify it in real time. The user interface generated in `liminal`
allows a user to play and pause the tour animation (via brushing),
panning and zooming to modify the scales of the plot to provide context and
click events to highlight groups of points (like clusters).

These interactions are enabled by treating the basis generation as a reactive
stream. Instead of realising the entire sequence, which limits the animation
to have a discrete number of frames, new bases and their intermediate frames
are generated dynamically via pushing the current projection to the visualisation interface. The interface listens to events like pressing a button
or mouse-dragging and reacts by pausing the stream. This process allows the
user to manipulate the tour in real time rather than having to fix the number
of bases ahead of time.
